---
title: "Class 6"
author: "Rachel Ng"
date: "January 23, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

## Sub Header 1

### Sub Header 2

plot(1:5, col = "red")

```{r}
# this is another plot in r
plot(1:5)
```

## More about file import
`read.table()` and friends for flat files
for example: read.table("File name", "Header = TRUE/FALSE, sep = " ")

```{r}
#Reading test 1
read.table("test1.txt", header = TRUE, sep = ",")
#or
read.csv("test1.txt")
#Reading test 2
read.table("test2.txt", header = TRUE, sep = "$")
#Reading test 3
read.table("test3.txt", header = FALSE)
```

Example function 1:
```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
add(x=1, y=4)
add(1, 4)
```
Let's use this functions on vectors
```{r}
add( c(1,2,4), c(1,2,4) )
```

What is this `range()` function?
```{r}
x<- c(4, 4, 10, 3, 11)
max(x)
min(x)
```

```{r}
range(x)
```

```{r}
rescale <- function(x) {
  rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale( c(1,2,NA,3,10) )
```

Let's improve rescale function to exclude NA
```{r}
rescale <- function(x) {
  rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}

#trying it again
rescale( c(1,2,NA,3,10) )
```

Rescale 3
```{r}

rescale3 <- function(x, na.rm = TRUE, plot = FALSE) {
  
  rng <-range(x, na.rm=na.rm, plot = FALSE)
  
  print("Hello")
  
  answer <- (x - rng[1]) / (rng[2] - rng[1])
  
  print("is it me you are looking for?")
  
  if(plot) {
    plot(answer, typ="b", lwd=4)
    }
 
  print("I can see it in ...")
 
  return(answer)
}

rescale3(1:10)
```

**SEE LECTURE 6 Hands-On worksheet for more about Range function**
1. read.pdb returns list of class "pdb" with 8 elements
file needs to be name of a PDB file
2. trim.pdb returns a list with 7 elements based on selections of atoms, in this case A chain alpha carbons "elety""
3. sse displays rectangles, which describe secondary strucutre of sequences


```{r}
library(bio3d)

s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor", helix.col = "red", sheet.col = "blue")s
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor", helix.col = "red", sheet.col = "blue")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor", helix.col = "red", sheet.col = "blue")

hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

4. **plotting on the same graph**
```{r}
plotb3(s1.b, sse=s1.chainA, lwd = 2, typ="l", col="black", ylab="Bfactor", helix.col = "red", sheet.col = "blue")
points(s2.b, typ="l", col="red", lwd=2)
points(s3.b, typ="l", col="green", lwd=1)

```



