---
title: "class10"
author: "Rachel Ng"
date: "2/6/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Review: 
PCA projects the features onto the principal components.
The motivation is to reduce the features dimensionality while
only losing a small amount of information.

#Unsupervised Learning: Human Breast Cancer Cells
```{r}
#load Wisconsin data

wisc.df <- read.csv("WisconsinCancer.csv")

View(head(wisc.df))

#convert to matrix
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)

#assigning rownames
row.names(wisc.data) <- wisc.df$id

diagnosis <- wisc.df$diagnosis

```

Q1. How many observations are in this dataset?
569 observations of 33 variables
Q2. How many of the observations have a malignant diagnosis?
212
Q3. How many variables/features in the data are suffixed with _mean?
```{r}
length(grep("_mean", colnames(wisc.data)))
```

##Principal Component Analysis
```{r}
#checking column means/SD
colMeans(wisc.data)

apply(wisc.data,2,sd)

#Perform PCA on wisc.data
wisc.pr <- prcomp(wisc.data, scale = T)

# Look at summary of results
summary(wisc.pr)


```

From your results, what proportion of the original variance is captured by the first principal components (PC1)?
44%
How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
Up to, including PC3
How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
Up to, including PC7

```{r}
biplot(wisc.pr)

# Scatter plot observations by components 1 and 2
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis, xlab = "PC1", ylab = "PC2")

#scatter plot observations by components 1 and 3
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis, xlab = "PC1", ylab = "PC3")
```

## Variance explained
```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)

#variance explained by each PC
pve <- pr.var / sum(wisc.pr$sdev^2)

plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

##Clustering
```{r}
wisc.pr.hclust <- hclust( dist(wisc.pr$x[,1:3]), method="ward.D2" )
plot(wisc.pr.hclust)

#groups with k = 2
grps2 <- cutree(wisc.pr.hclust, k =2)
table(grps2)

plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps2, xlab = "PC1", ylab = "PC2")

View(table(grps2, diagnosis))
```
Red corresponds to benign and black corresonds to malignant

```{r}
g <- as.factor(grps2)
levels(g)

g<- relevel(g,2)
levels(g)

plot(wisc.pr$x[,1:2], col=g)
```

##sensitivity/specficity
sensitivity = test's ability to correctly detect ill patients who do have the condition
specificity = test's ability to correctly reject healthy patients w/o a condition

url <- "new.samples.cs"
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc

#plot
plot(wisc.pr$x[,1:2], col = g)
points(npc[,1], npc[,2], col = "blue", pch = 16, cex = 3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

Q17. Which of these new patients should we prioritize for follow up based on your results?
Patient 1!