---
title: "class13handson"
author: "Rachel Ng"
date: "February 18, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bio3d)
file.name <- get.pdb("1HSG")
```

```{r}
hiv <- read.pdb(file.name)
resid <- atom.select(hiv, "protein", value=FALSE)
```

Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 
A. Non protein/nucleic resid. values are: HOH, and MK1
B. use of `atom.select()`
resid <- atom.select(hiv, "protein", value=FALSE)

#Prepare initial protein and ligand input files
first have to extract just the protein atoms...

```{r}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

#writing out new pdb files
write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")
```

#Using AutoDockTools to setup protein docking input
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi = T)
write.pdb(res, "results.pdb")
```
Using `rmsd()` to compare ligand function
```{r}
#res <- read.pdb("all.pdbqt", multi = TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)

```


